### 2023-02-06 draft.Rmd
modifies script to add what Kevin wrote and uploads initial data

---
title: "Extended early adulthoods"
author: "Kevin Kiley & Nicolas Restrepo"
fontsize: 12pt
output:
  pdf_document: default
  html_document: default
header-includes:
  - \usepackage{setspace}
  - \doublespacing
  - \usepackage{lineno}
  - \linenumbers
  - \def\linenumberfont{\normalfont\tiny\sffamily}
bibliography: efw_bib.bib
abstract: "Young adulthood is consistently identified as a “formative period” in the life course, when people have heightened probability of attitude change. This probability has often been attributed to the social conditions of adolescence. Over the past 70 years, changes in the timing of major life course events such as marriage have transformed and extended this life stage. We leverage these changes to explore whether these life course changes, as well as broader social changes, have affected how likely people are to make durable changes of attitudes. We find strikingly little difference in the rates of any kind of stable opinions and opinion change among young people in different eras. People under 30 are similarly different from people over 30 in all eras. The overall pattern suggests either that social conditions are not strongly related to attitude stability or that different social shifts have had counter-vailing effects that produce aggregate stability in rates." 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Adolescence and young adulthood are typically identified in social science literature as “impressionable years” – a stage of the life course in which people are more susceptible to broad social currents, when their opinions are more malleable, and when peoples’ core ideas and attitudes are formed []. Research regularly finds that people are more likely to change their attitudes and opinions in these life stages than other life stages, with opinions becoming more stable once these life stages have passed []. This pattern, in which young people are open to attitude change and become resistant over time, underpins dynamics around cohortization that are frequently invoked in larger models of social change [@ryder1962; @mannheiumYEAR].

However, why people in these life stages appear to show more frequent opinion change relative to older people is unclear. Work in biology, psychology, and cultural evolution shows that peoples’ cognition changes in ways that might make them more resistant to new information as they age []. Explanations in social sciences, in contrast, tend to highlight the unique social features of these life stages – high rates of mobility between social contexts, participation in educational institutions, and cultural expectations about pursuing “new experiences” – that shape how people are exposed to new sources of information or their ability to update opinions in the face of new information []. While both mechanisms likely play a role in facilitating opinion change in this life stage, the relative contributions of each are unclear.

Clarifying the mechanism that drives the frequency of opinion change during this formative period is important for understanding the broader process of social change and cultural evolution. If openness to change in early adulthood is principally a function of the social structure of life in that time period, changes in the timing and ubiquity of life course events like marriage and childbearing can have significant impacts on the broader process of macro-social change []. Similarly, social changes such as the deinstitutionalization of marriage [@cherlin2004; @cherlin2020; @robbins2022] that decrease distinctions between life course stages might make older people more open to change in more recent eras than in older eras, diminishing the importance of “cohorts” for social change. If the mechanisms are principally cognitive and biological, rather than social, then these changes are unlikely to affect rates of social change in the same way.

In this paper, we seek to further elaborate this debate by exploring how changes in the broader social structure and changes in the timing of life course events since the 1950s have affected rates of change in political attitudes over time, especially differences in rates of opinion change between young adults and the rest of the population. Over the course of the 20th century, the timing of life-course transitions such as completing schooling, entering the workforce, getting married, leaving one’s parents’ home, and having children, have changed dramatically, with most of these event occurring later in the life course in more recent cohorts [@buchmann; @bruckner2005]. Similarly, this stage of life has become more detached from family and other role obligations, giving people more freedom to explore new ideas and lifestyle choices [@rosenfeld2007]. This suggests both a lengthening and a intensification of this stage of the life course in ways that should increase the rates at which people make changes of opinion.

We take advantage of this change over time to explore whether young adults of different cohorts demonstrate more or less change in general political opinions over time relative to other members of the population. We compare rates of opinion change in comparable general political attitudes across four eras – 1956-1960; 1972-1976; 1992-1996; and 2006-2010 – and between young people and old people in these eras to evaluate a range of competing hypotheses about what accounts for the divergence in attitude behaviors over the life course.

Broadly, we find that the probability of political opinion change in people under 30 have not increased as the “early adulthood” window has lengthened and the conditions of this life stage have changed. Across eras, young people make political opinion changes at remarkably similar rates, with social transitions such as marriage, having children, and completing education playing little role in the rates at which these opinion changes happen. While these findings do not provide conclusive evidence of one mechanism over the other, they confine the range of potential theories of attitude development and the nature of political opinions. In particular, they suggest that there is something relatively timeless – potentially biological and cognitive factors – that accounts for increased opinion variance in young people.

## Impressionable Years / Opinion Change and Stability

The idea that the early stages of the life course are especially influential in the process of attitudinal formation is dominant across contemporary accounts of socialization (Dinas 2014; @guhin2021). The argument – at times implicit – is that the circumstances in which people are brought up leave an enduring and traceable mark on their worldviews (Gerber and Green 1998). In the study of political socialization, this perspective enjoys a longstanding history. “Generational effects” are used to explain the development of political preferences, longitudinal shifts in attitudes, and differences across age-groups (Mannheim 1970; Bartels and Jackman 2014). However, the notion of “impressionable years” is more general. The burgeoning literature on cohort effects rests on the assumption that our early years play key role in shaping not just political opinions, but peoples’ cognition on many dimensions (Elder 2018). Their early years will leave a mark that will be predictive of their life outcomes, well beyond those impressionable years have passed. Thus, the notion of a formative period – at the early stages of the life-course – underpins diverse and longstanding lines of research.

The “impressionable years” hypothesis – that people demonstrate more opinion change early in life followed by a period of relative stability, especially in middle age – has often been tested against other theories of opinion change over the life course. The most common are an aging stability or increasing persistence hypothesis that argues that openness to change gradually declines over time and a life stages hypothesis, in which people are open to attitude change both early in life and later in life, with stability principally happening in middle ages [@visserkrosnick1998].  While different attitudes occasionally follow different patterns, the impressionable years hypothesis has generally been shown to be true for broad political ideology, partisan identification, and specific policy preferences [@krosnickalwin1989; @alwinkrosnick1991; @searsfunk1999]; social attitudes [@danigelis2007]; and interest in politics [@prior2010]; as well as things like personality [@wagner2019]. More generally, recent research has found that, with a couple notable exceptions of high-salience events such as national shifts around gay rights, durable change of social opinion is quite rare among U.S. adults [@kileyvaisey], and Vaisey and Lizardo [-@vaiseylizardo2016] find that cohort explains much more variance than period for most items in the General Social Survey, suggesting a strong impressionable years pattern.

In recent years, researchers have attempted to use new approaches to quantify “openness” to social influence across the life course, especially around politics [@bartelsjackmanYEAR; @ghitzagelman]. This work has produced remarkably similar patterns with openness peaking from a person’s mid-teens through age 25 and declining quickly after that, lending more credence to the “impressionable years” hypothesis. However, this work tends to assume a constant age-based rate of openness across cohorts – that 20-year-olds in 1950 are open to influence at the same rate as 20-year-olds in 2010 – which has not been tested.

## Theoretical Models

Why is it research consistently finds the period of time between 14 and 25 to be distinctly “impressionable years?” At the outset, it is important to note that changing one’s mind involves at least two distinct processes that can be difficult to disentangle: exposure to new ideas and openness to change conditional on exposure to new ideas. For example, it could be the case that people of all ages are exposed to new ideas at similar rates, with acceptance of a new idea being a function of age, which would account for age-based patterns in opinion change. Conversely, it could be the case that rates of exposure to new ideas changes over the life course, while openness to changing one’s mind as a function of exposure does not, producing the same pattern. It could also be the case that both change as a function of age. Any patterns we observe should be interpreted as coming from changes in one or both of these processes. In this section we outline theoretical mechanisms that have been suggested to contribute to one or both of these processes.

### Biological Mechanisms

The first potential mechanism for changes in attitude stability over time is biological: that the development of the of human brain as people age makes it more difficult for people to make changes of opinions after some point in their life. This approach tends to suggest that human cognition works similar to Bayesian updating (Bartels and Jackman 2014; Sherratt and Morand-Ferron 2018; Gopnik 2020). In this framework, humans weight information equally, whether they encounter it during adolescence or middle age. However, given that by the time people reach adulthood they have already encountered a lot of information, each new piece will have less of an impact (Bartels and Jackman 2014). The distinctive thing about our early years is simply that they happen early. They prove so influential because the incremental effect of new information diminishes as we age.

It is also possible that there is something unique about human cognition during these periods: that agents ascribe more weight to the information acquired within them. Maturational arguments about the increased plasticity during our early years are consistent with this position (Mata and von Helversen 2015). Our brains exhibit more malleability during childhood and adolescence (Mata and von Helversen 2015). This is consistent with the fact that these early formative periods have been observed across species and contexts. It is not only human children that exhibit a heightened capacity to adapt. In a study of social learning among birds, for instance, Aplin et al (2017) show that younger individuals – unlike their older counterparts – were consistently able to retool their habits in light of new information.

Though the Bayesian updating explanation is simple and elegant, research across different disciplines has produced evidence that supports the existence of these periods of heightened sensitivity. In light of empirical evidence about these sensitive windows, there has been extensive modeling work about their evolution. This line of research seeks to understand why periods of heightened sensitivity would emerge and what contexts would facilitate their emergence (Fawcett and Frankenhuis 2015). The central argument is that the emergence and timing of these sensitive windows depends on the interaction between the access to information, the reliability of information, and the stability of the environment (Frankenhuis and Walasek 2020). Early formative periods are likely to evolve in specific configurations: when environments are unstable, and information is plentiful yet unreliable (Frankenhuis and Walasek 2020). In these circumstances, it makes sense for agents to explore early on, to find something that works and then to stick to strategies that have worked well in the past. Access to different types of information might vary across time and this might explain the emergence of sensitive periods at later stages of the life-course (Walasek, Frankenhuis, and Panchanathan 2021). For example, the importance of musical tastes and fashion choices becomes important around adolescence, and, at this stage, we are awash with – at best contradictory – information about what one should or should not like, should or should not wear. Sensitive periods around the development of these tastes might peak around this period, when there are more signals and when the payoffs are more salient.

The thrust of the biological arguments is that, through millennia of evolution, humans have developed a cognitive architecture that makes them relatively open changing dispositions at younger ages and relatively resistant at older ages. However, while biological mechanisms can shape how individuals respond to new information when they encounter it, these processes cannot dictate the rate at which individuals are exposed to new information, which might also matter for rates of opinion change. For that, we must turn attention to the structuring of the social world.

### Life Course // Social Role Mechanisms:

Humans’ early years might be distinct in terms of cognitive plasticity, but they are also socially distinct: they are characterized by sustained social learning and punctuated with constant changes of social contexts. Here we briefly consider some of the ways that the “impressionable years” differ in ways that might expose people to new information at higher rates and also affect their willingness to update beliefs, beyond what cognition might allow.

First, people in this age group tend to move through social contexts much more quickly than people in other age groups. In the 11-year window from ages 14 to 25, in which people seem to show the highest rates of opinion updating, young people move from middle school to high school to either higher education or the workforce or both, with each year in these contexts potentially exposing people to new classmates, teachers, and other role models, peers, and sources of information. People in this age window also show the highest rates of residential mobility, moving for educational and occupational reasons [is this true?]. Assuming that new social contexts shape exposure to new information, this heightened contextual mobility during this stage might explain part of the difference between age groups.

Second, people in the “impressionable years” window spend their time in different social contexts than people in middle age. Data from the American Time Use Study shows that people under age 25 spend a larger portion of the day on leisure time and educational activities and a smaller part of the day doing work and work-related activities than people between the ages of 25 and 55. These contexts, especially education, are designed to expose people to new information in ways that can lead to new opinions. For example, there is some evidence that college attendance leads to changes in opinions about civil liberties, gender egalitarianism, and racial inequality [@campbell2016; wodtke2018]. Social networks also differ in this time period. Previous work has found that young peoples’ networks tend to be larger and include more friends than kin, for example [@marsden1987].  It could be the case that larger “weak tie” networks expose younger people to more new information that drives opinion change [@granovetter1975; burt2005].

Third, these are periods in which exploration and self-discovery are expected and encouraged; something that should not be taken for granted given social variation in norms around when maturity is reached [@cunningham2012; @elder1994].

### Updating conditional on exposure

While social conditions might expose young people to new information at higher rates, it might also make it easier for them to act on these new sources of information in ways that facilitate opinion change.

Previous work on stability of attitudes over the life course suggests that people tend to value stability in figures of authority. As people move into roles with more authority in mid-life, both in family and occupational contexts, they feel this expectation from others and act accordingly, closing themselves off to updating beliefs [@eaton2009]. This potentially explains another phenomenon with some evidence behind it: that people appear to open up to changing their minds again as they enter post-retirement [@danigelis]. At this point, peoples’ authority diminishes, potentially opening them up to change later in life.

Similarly, mid-life is associated with more homogenous networks [@marsden1985; @diprete2011], which experimental research suggests reduces the propensity to adopt new beliefs [@vissermirible]. If young people exist in less dense, more heterogeneous social networks, they might it easier to update their networks to reflect new beliefs rather than constrain their beliefs to their social networks [], or close off parts of their lives to people who might not be supportive of new attitudes.

Given the distinct social conditions of the early-adult phase of life, there are significant reasons to believe that the “impressionable years ” are principally an effect of how we have chosen to structure the life course, or the interaction between these life course stages and biological mechanisms.

## Changing Life Course

Contemporary historical trends offer the opportunity to disentangle the role of socio-cultural and  biological factors in explaining why we observe “impressionable years.” In the U.S., the last decades have seen considerable changes around adolescence and young adulthood (Buchmann 1989; Rosenfeld 2007; Rosenfeld and Kim 2005). The expansion of higher education and increased possibilities for mobility are among the many social processes that have led to changing expectations for the trajectories that young adults should follow (Rosenfeld 2007). In general, as Buchmann  (1989) notes, life trajectories have become less predictable. The “script” of young adulthood has become less rigid: it might involve prolonged higher education, moving to a different city, or dating without a stable romantic partner. The routes that people are taking into adulthood are increasingly diverse. For the purposes of this paper, there are two accompanying trends that are particularly important: the elongation and intensification of young adulthood.

### Timing

The period between adolescence and “settling down” seems to be getting longer, and this might affect the patterns of attitudinal updating we expect to see for the same age groups in different cohorts. Buchmann (1989) notes that the diversification of life-trajectories has been accompanied by the postponement of the rituals associated with adulthood. In the U.S., the median age at first marriage has consistently risen in the last decades and so has the median age at which families are having their first child (Rosenfeld 2007). We also see higher rates of non-marital childbearing across all demographic groups [@cherlin]. Thus, the markers of the transition into adulthood seem to be occurring later in the life-course and potentially in different orders. Therefore, the definition of what counts as our “impressionable years” might be extending. This, in turn, could have an effect in how long exploration lasts throughout the life-course. If this is the case, we should expect to see more attitudinal updating among individuals in age-ranges that, in previous cohorts, had displayed mainly stability. In other words, people who were born say in the mid 20th century would have displayed mainly attitudinal stability by the time they reached their mid-twenties while more contemporary counterparts might still exhibit attitudinal updating when they reach that age. Adolescents are taking longer to move through the rituals of adulthood, and this might in turn elongate the period in which exploration and self-discovery are expected.

### Changing Impressionable Years

Another second trend relates to just how much exploration young adults can engage in during these “impressionable years.” In his work, Rosenfeld (2007) explores the rapid rise in same-sex and interracial marriages in the U.S. His argument is that one of the processes underlying these changes is that young adults are increasingly escaping the controlling gaze of their parents. When young people live with – and/or financially dependent on – their parents, the latter have more say about their children’s choice of partner. The increased independence and mobility that young adults have enjoyed since the mid 20th century means that they have been able to engage in courtship outside the gaze of their parents, thus making choices that would not have been sanctioned by them. Rosenfeld’s (2007) work highlights an important insight: exploration might not be only a product of how sensitive we are to novel information but also how – and whether – we can act upon such information. The social institutions where we spend our impressionable years might shape how much exploration we can engage in. An increase in independence then might entail more exploration. This assertion in turn engenders a prediction: we should see more attitudinal updating among contemporary young adults than their older counterparts, who tended to be closer to the grip of familial control, even if they have met similar life course milestones. The contemporary social organization of young adulthood, which has been developing for the last few decades, might give rise to more meaningful attitudinal changes by providing more leeway to deviate from established social norms.

This approach suggests that it is not the timing of life events that is important for rates of change, but the broader social conditions of early adulthood that might make people more malleable in recent cohorts even net of changes. In other words, unmarried, childless 25-year-old men and women in 1956 might have experienced a set of social constraints that unmarried, childless 25-year-old men and women in 2008 do not experience, leaving the latter much more open to developing new opinions than their predecessors.

## Hypotheses

Sociological work then highlights that there are principled reasons to believe that length and intensity of formative periods might be affected by social factors, and yet this question remains unexplored. The importance of youth remains the backbone of sociological theories of socialization and there has been work trying to outline the cognitive assumptions that would lead to marked influence of our early years. While the evidence for these formative periods has accrued, there has been no research on the social variability of these periods. Any evidence of differences in the rate of attitudinal updating among people of similar ages across different social contexts would suggest that sensitive periods are not only maturational or a result of “tallying up” information. In turn, finding similar patterns of attitudinal updating across widely different social contexts would lend further credence to the idea that our early years are “impressionable” either because we weight information more heavily during these periods or because we are cementing our opinions, regardless of the social contexts we happen to find ourselves in.

The argument so far implies three hypotheses that we will test in our study. Given historical changes in the meaning of adolescence, we hypothesize that we will see differences in the rates of attitudinal updating that the same age-groups across different cohorts will exhibit. The first hypothesis then is non-directional:

H1: There will be differences in the patterns of attitudinal updating from individuals in the same age groups but in different cohorts.

The second hypothesis is based on Buchmann’s work and it concerns the fact that young adults are increasingly postponing the rituals associated with the transition into adulthood. Thus, we posit that more recent cohorts should display more exploration well into their twenties than previous cohorts.

H2: Young adults in recent cohorts should exhibit active attitudinal updating later than their counterparts of similar ages from older cohorts.

The last hypothesis in turn relates to Rosenfeld’s ideas. Young adults in recent cohorts have enjoyed more independence and, therefore, might have had more opportunities to substantially change their beliefs and habits. We hypothesize then that:

H3: Young adults in recent cohorts should exhibit higher rates of attitudinal updating than people of a similar age in previous generations.

Finally, it could be case that an explanation for overall change in rates of the population are explained by factors that we have not explicitly considered here. It could be the case that rates of opinion change are tied to things like racial and ethnic identity, religious identity, or geographic region, all of which have changed significantly in the same window. If we identify broad shifts in attitude change that are not explained by , this suggests .

## Kevin's Hypotheses

Kevin’s Hypotheses:
-	Broadly speaking the “impressionable years” hypothesis would suggest that: People under 30 will demonstrate higher rates of attitude change than people over 30 across all cohorts.
-	Both the “extended adolescence” and the “intensification of young adulthood” theories would both suggest that people under 30 will demonstrate higher rates of attitude change (relative to people over 30) in more recent panels than people in earlier panels. This could be because people 25-30 make changes at higher rates than they were in the past or because people 18-25 and 25-30 make changes of attitudes at higher rates.
-	Adjudicating these might require separating people 18-25 from 25-3X. But this seems tough since the numbers are too small in most of the panels to separate these. But I might try.
-	The “timing of life course” events hypothesis would seem to suggest that: “Controlling” for marital status and having children [and maybe other stuff] will reduce the difference between people under 30 and people over 30. 
o	In other words, net of age, we should see that marriage and having children makes people more stable opinion holders (maybe also something about education). So I’m going to try to include these variables in some analyses. 
-	The “intensity of early adulthood” argument would seem to suggest that: Controlling for differences in life course stages, younger people in more recent panels will still demonstrate higher rates of attitude change than younger people in older cohorts.

## Kevin’s Other Thoughts
-	There’s the possibility that people over 30 have increased their likelihood of making a change. First, the social environment is significantly different: people are likely exposed to new information at higher rates. Second, deinstitutionalization of the life course and increases in divorce rates, etc. might mean that older people are more free to make opinion changes than they were in the past too.
-	Similarly, there’s the “life stages” hypothesis (its not that early life is distinctly open, it’s that middle age is distinctly closed) that has some support. It might be worth either holding out or including a separate coefficient for people > 60 or something. 
-	There’s a possibility that we see differences in rates of change between cohorts that are not explained by life course stuff. One thing that I’ve been thinking is like a “population composition” thing. Black and Hispanic people are less stable opinion holders than white people on a number of issues (something I’ve been thinking about in my spare time). People who attend religious services are also more stable opinion holder than other people. Changes in this stuff might be worth considering. Or maybe it’s part of the discussion of the results.

## Data

To test the hypotheses outlined above, we require data on within-individual change in comparable attitudes over a stable window of time across eras. To do this, we focus on general political beliefs, including views on the role of government, political ideology, and partisan identification, that have been asked in panel surveys. This is in part a practical matter, as these are the kinds of questions that have been included in panel studies that date back to the mid-1950s, giving us the largest window of cohorts to explore. However, these kinds of questions have frequently been used to identify the “impressionable years” [].

We draw on data from four different survey panels: the 1956-60, 1972-76, and 1992-96 American National Election Study panels , and the 2006-10 General Social Survey Panels. All four surveys are nationally representative samples of non-institutionalized adults in the United States. All four panels survey respondents three times over a four-year window, making the durations comparable.

### Political Opinions

We sought to identify questions that tapped general political sentiments that were echoed in at least one other survey. We avoided questions that asked about specific policies or political figures, since they were often not replicated across panels. We also focused principally on questions asking about the role of the federal government in domestic affairs, rather than international concerns.

No questions are directly comparable across each of the panels. For this reason, we include a broad range of questions, in hopes of establishing general patterns. Table XXX outlines the questions used in this analysis, their specific wording, and response options.

### Life Course Events

To evaluate changes in the life course, and their relationship to attitude stability, we focus on a set of events that are traditionally accepted as marking the transition to adulthood.

While question wording differs slightly across surveys, all surveys include a measure of whether people are married, whether they have any children, and .

For example, while marriage rates have declined , rates of cohabitation (romantic partners living together without being married) have increased.

## Analytical Strategy

It is not obvious what it means for a person to change their opinion in a survey, and debates about how to measure “change” in opinions often focuses on differences in the approach to measurement [@converse; @achen1975; ]. Rather than adopt a single approach to quantify the rates of change across questions, we use a variety of methods with different assumptions about what “change” entails and attempt to triangulate the findings to see if patterns are robust across these assumptions.

The first approach to measuring change in political opinions is to assume that all change is meaningful and measure the total “distance” people move across a scale during the time period in question.

A second approach is to a prioi define what kinds of opinion behavior count as “change” and define what counts as “stable opinions” and compare their prevalence across age groups and panels. For example, we might count any response pattern in which people stay on the same side of a midpoint (“agree”-“strongly agree”-“agree”) as a stable opinion and any response option that crosses the midpoint (“agree”-“disagree”-“disagree”) as representing a change of opinion.

A third and related approach is to classify people who change their responses within a specified range, such as changing less than two scale points (e.g., “agree”-“neither”-“agree”), as expressing stability while counting other respondents as making changes of opinion.