### 2023-08-11 Analyses/extended_abstract_asa23.Rmd
edits 811

---
title: "Extended Abstract"
author: "Kevin Kiley & Nicolas Restrepo"
fontsize: 12pt
output:
  pdf_document: default
  html_document: default
header-includes:
  - \usepackage{setspace}
  - \doublespacing
  - \usepackage{lineno}
  - \linenumbers
  - \def\linenumberfont{\normalfont\tiny\sffamily}
bibliography: ./efw_bib.bib

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F)
library(tidyverse)
library(brms)
theme_set(jtools::theme_apa())
```

## Motivation

Adolescence and young adulthood are typically identified in the social science literature as "impressionable years" -- a stage of the life course in which people are more susceptible to broad social currents; when their opinions are more malleable; and when their core dispositions, identities, and opinions are formed [@alwin1991]. A broad body of research finds that rates of opinion change tend to peak during the window of ages 14-25, with opinions becoming more stable once people move into middle age [@sears1999; @bartels2014; @ghitza2022]. And this recurring empirical pattern is central to broader dynamics around cohortization that are frequently invoked in larger models of social change [@ryder1965; @mannheim1952; @vaisey2016].

However, why people in these life stages appear to show more frequent opinion change relative to older people is unclear. Work in biology, psychology, and cultural evolution suggests that peoples' cognition changes in ways that might make them more resistant to new information as they age. Explanations in sociology, in contrast, tend to focus on the unique social features of these life stages -- including high rates of mobility between social contexts and participation in educational institutions -- that shape how people are exposed to new sources of information and their ability to update opinions in the face of this information [@visser2004a]. While both mechanisms likely play a role in explaining the higher rates of opinion change in this life stage, the relative contributions of each are unclear.

In this paper, we leverage changes in the timing, ubiquity, and standardization of life course transitions over the second half of the 20th Century [@buchmann1989; @bruckner2005] to adjudicate the relative contributions of developmental and social mechanisms to the heightened rates of opinion change we observe in young people. Since the 1950s, the median age of marriage and childbirth have each been pushed back by about 10 years for both men and women [@eickmeyer2017; @manning2014], rates of college-going have increased dramatically, and people make more transitions between workplaces and between the workforce and education [@bruckner2005]. These shifts have led to a dramatically expanded window of participation in educational institutions, geographic and social mobility, and freedom from various social obligations and expectations, all of which are assumed to facilitate opinion change [@rosenfeld2009]. If the heightened rates of opinion change we observe in young adults are principally due to the social structure of this life stage, then we should observe changes in the rates of opinion change in this group over time.

Drawing on four nationally representative panel surveys, dating from the mid-1950s through the early 2000s, we compare multiple different measures of opinion change in general political opinions and partisan identification of early adults (people under 30) and the rest of the population. Despite the significant changes in the standardization of the life course and the duration of early adulthood, we find remarkable consistency in opinion behaviors over time, both for young adults and for the rest of the population. Across multiple measures of opinion change, we find that people under 30 change their opinions at slightly higher rates than people over 30, but this gap remains stable across panels. The overall pattern of results suggests either that the heightened rates of attitude change in early adulthood are principally driven by relatively constant developmental factors -- how the brain reacts to new information as people age -- or that the dramatic changes in the social structuring of early adulthood have combined in a way to produce aggregate stability over six decades.

## Data & Analytic Strategy

We draw on data from four different panel surveys of the American population: the 1956-60, 1972-76, and 1992-96 American National Election Study panels, and the 2006-10 General Social Survey panel. All four surveys are nationally representative samples of non-institutionalized adults in the United States. All panels also interviewed survey respondents three times over a four-year window, making the duration of each survey window comparable. All four panels also include comparable questions about general sentiments toward government action, a measure of partisan identification, and measures of life course transitions such as marriage and childbearing.

The broadest challenge for our analysis is that it is not obvious what it means for a person to demonstrate opinion change in a survey. Thus, rather than adopt a single approach to quantify the rates of change across questions, we use a variety of approaches. We calculate whether participants changed at all across the three waves (), their absolute movement across the scales (), whether they crossed the midpoint of the scale at any point (), and whether their last position is at least two points away from their starting one (). All these outcome variables presuppose different assumptions about what "change" entails and, thus, we present them all in an attempt to identify patterns that are robust across a range of assumptions.

For every definition of change, we regress the outcome measure on an indicator variable for whether people are aged 30 or less. For the continuous outcomes, we use ordinary least squares regression and for the dichotomous outcomes we use logistic regression. This produces an extremely large number of models for each of many questions, and will likely produce statistically significant coefficients at the $p < .05$ level for a substantial proportion of coefficients simply by chance. Our goal is not to interpret any single coefficient but to interpret overall patterns if they emerge.

## Results

Figure 1 plots coefficient estimates generated from regression models of change on whether respondents are under 30. Each column represents a different measure of "change" outlined above. ... If young people have become more likely to make

```{r}

```

Figure 1 plots predicted differences between young and old people generated from the regression models outlined above. These are presented in log-odds for the dichotomous outcomes and scalar differences for the continuous (total change/standard deviation) outcomes. If rates of attitudinal updating among young people had been increasing across time, we would see the plotted coefficients becoming larger as panels become more recent. Here, it means that the brown coefficients should be smaller than the blue/green ones, and the darkest blue/green one -- which represents our most recent panel -- should be the largest. Instead, in general, there are very few coefficients whose confidence intervals do not overlap 0, suggesting that young peoples' opinion behavior is not substantially different than that of older people for the issues examined here across all eras.

This is not to say that young people do not have substantively different opinions (they likely do), just that they are not any more or less stable by the metrics we outlined above. There are only a handful of issues where we observe a consistent pattern across most measures of attitude change: partisan identification in most panels, the issue of busing in the 1970s, and general views on government spending in the 1990 panel.

This is a rather striking pattern that challenges the notion that

## Discussion

Despite significant shifts in the timing of life course events and the broad social structure of early adulthood from the mid-1950s to today, there is little change in the rates at which people between the ages of 18 and 30 undergo major changes of political opinions. The patterns observed in this analysis broadly suggest that there is something developmental that explains broad patterns in attitude instability early in the life course. It could also suggest that the features that explain stability are still social in nature, but are rooted in factors that have not undergone significant changes since the 1950s, though it is unclear what those might be.